---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rquest

<!-- badges: start -->
[![R-CMD-check](https://github.com/shenal-dkumara/rquest/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/shenal-dkumara/rquest/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/shenal-dkumara/rquest/graph/badge.svg)](https://app.codecov.io/gh/shenal-dkumara/rquest)
<!-- badges: end -->

## Overview

The `rquest` package provides convenient functionality for researchers to carry out hypothesis tests and obtain confidence intervals for measures based on quantiles. This includes for single quantiles (e.g., the median), linear combinations of quantiles (such as the interquartile range), ratios
of linear combinations commonly found in skewness and kurtosis measures, and newly developed inequality measures. Another key objective is to
make it easy for users to define their own measures for hypothesis testing and confidence intervals.

Following are the main functions in the package:

* `q.test()` carry out hypothesis tests and obtain associated confidence intervals for linear combinations of quantiles, and ratios of such linear combinations.

* `qineq()` carry out hypothesis tests and obtain associated confidence intervals for quantile based inequality measures

* `qcov()` compute a covariance matrix consisting of variances (on the diagonal) for quantile estimates and covariances (off-diagonal) between different quantile estimates


## Installation

You can install the development version of rquest from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("shenal-dkumara/rquest")
```


## Usage

```{r example}
library(rquest)

## Functionality of q.test() ##

#  Create some data
x <- c(8.43,7.08,8.79,8.88,7.87,5.94,8.79,5.46,8.11,7.08)
y <- c(13.44,13.65,14.77,9.51,14.07,10.92,11.59,13.42,8.93,10.88)

# One sample hypothesis test for the IQR
q.test(x, measure = "iqr")

# Two samples hypothesis test for robust coefficient variations (0.75*IQR/median) with log transformation and back-transformation to the ratio scale,.
q.test(x, y, measure = "rCViqr", log.transf = TRUE, back.transf = TRUE)

# The same two samples hypothesis test for robust coefficient variations (0.75*IQR/median) by using 'u',''u2','coef' and 'coef2' arguments.
u<-c(0.25,0.75)
coef<-0.75*c(-1,1)
u2<-0.5
coef2<-1
q.test(x,y,u=u,u2=u2,coef=coef,coef2=coef2,log.transf=TRUE,back.transf=TRUE)

# The same two samples hypothesis test for robust coefficient variations (0.75*IQR/median) by using only 'u' and 'coef' arguments.
u<-c(0.25,0.5,0.75)
num <- 0.75*c(-1,0,1)
den <- c(0,1,0)
coef <- rbind(num, den)
q.test(x,y,u=u,coef=coef,log.transf=TRUE,back.transf=TRUE)

## Functionality of qcov() ##

# Compute the variance-covariance matrix for sample quartiles.
qcov(x, c(0.25, 0.5, 0.75))


## Functionality of qineq() ##

# Two sample hypothesis test for the QRI measure
qineq(x,y)
```




